---
title: "AAO2021_RExercises"
author: "Naweed Chowdhury, Rijul Kshirsagar, Jonathan Liang"
date: "10/6/2021"
output:   
  pdf_document: default
  html_document: default
---

*R for Data Analysis:  Useful Tools for Otolaryngologists and Researchers*

 \vspace{10pt}
 
**Instructions:**

Download the 'CRSData' dataset from the the shared google drive folder (https://drive.google.com/drive/folders/1pe0im70ZLKa2wr_Jjp0McJmFmUEq31JD?usp=sharing).  *Note this is in csv format.*  Add this file into your working directory for R.

 \vspace{10pt}


**EXERCISES:**

 \vspace{10pt}

***PART A:  R BASICS***

 \vspace{10pt}

1.  Install the 'tidyverse' pacakge.  After you install, load the 'tidyverse' library. 
```{r, Q1, include = FALSE}
#Insert code here
install.packages("tidyverse") #this may take a few seconds to install all the tidyverse packages
library("tidyverse")
```

2. Check your working directory.
```{r, Q2, include = FALSE}
#Insert code here
getwd()

#You can set the working directory by going to Session --> Set Working Directory --> Choose Directory.
```

3. Make sure the 'CRSData.csv' is within your working directory folder.  Load the 'CRSData.csv' file into your R environment.  Name it 'CRS'.
```{r, Q3}
#Insert code here
CRS <- read.csv("CRSData.csv")

#You can also click on 'CRS' in the global environment (top right pane) and it will open in your source pane for you to view the dataset more easily.
```

4. Check the structure of the CRS data. 
```{r, Q4}
#Insert code here
str(CRS)

#Note some are 'chr' (characters) - categorical variables, while others are 'num' (numeric) or 'int' (integer) - continuous variables. 
```

5. Return the first few lines/observations in the CRS data.
```{r, Q5}
#Insert code here
head(CRS)
```

6. Check the column names in the CRS data.
```{r, Q6}
#Insert code here
colnames(CRS)
```

7. Provide a summary of the 'Phenotype' (i.e. how many patients have each phenotype?)
```{r, Q7}
#Insert code here
table(CRS$Phenotype)
```

8. Provide a summary of the 'Phenotype' stratified by 'Asthma' status (No, Yes).
```{r, Q8}
#Insert code here
table(CRS$Phenotype, CRS$Asthma)
```

9. Provide a summary of the 'PreSNOT' and 'PostSNOT' score (i.e. what is the min/max/mean/median of the PreSNOT and PostSNOT scores)
```{r, Q9}
#Insert code here
summary(CRS$PreSNOT)
summary(CRS$PostSNOT)
```

10. Calculate the difference in SNOT scores (PostSNOT - PreSNOT) and call this new variable 'deltaSNOT'.  Provide a summary of 'deltaSNOT'.
```{r, Q10}
#Insert code here
CRS$deltaSNOT <- CRS$PostSNOT - CRS$PreSNOT
summary(CRS$deltaSNOT)
```

 \vspace{10pt}
 
***PART B:  VISUALIZATIONS***

 \vspace{10pt}



 \vspace{10pt}
 
 ***PART C:  STATISTICAL ANALYSIS***

 \vspace{10pt}

